A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-18 FMCAD2021/5_optimization.tex
optimization

Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ and  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$. Eventually, we obtain a {\pa} formula.

\paragraph{Perform the elimination of exponential occurrences of variables in parallel on atomic formulas}

Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namingly $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters. Since the parameters are used as boundary conditions in enumeration, we can choose the strictest boundary conditions and apply to all atomic formulas.

@@

For instance, assume coefficients of $10^\ivarx$ in atomic formulas of $\varphi$ are all non-negative. Let $\varphi^+$ denote the formula obtained by replace atomic formulas where coefficients of $10^\ivarx$ are positive by $\ltrue$. The resulted formula $\varphi'$ is
\[
\small
\begin{array}{l}
\bigvee \limits_{p=0}^{B^{\max}-1} ( \ivarx_{\sigma(i-1)} = p \wedge \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) [p/\ivarx_{\sigma(i-1)}])   \\
\bigvee \left(
    \begin{array}{l}
        \ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  - 1  \wedge \\
        \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} +\delta^{\max} \wedge \varphi^+  
    \end{array}
    \right)\\
%
\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  -1 \ \wedge \\
 \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right)\\
\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \ge \alpha^{\max}_2 + 2 \ \wedge \\
 \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right)\\
\bigvee \bigvee \limits_{p = \alpha^{\min}_1}^{\alpha^{\max}_2 + 1}
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} = p\ \wedge \\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right),
\end{array}
\]

@@

The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size. If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory. Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts. If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' result is reported.

\paragraph{Preprocessing with small upper bound}

>>>>

Similarly, we can eliminate the exponential occurrences of $\ivarx_{\sigma(i-2)}$ from $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}]$ as well as  $\tau'(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}]$, and so on. Eventually, we obtain a {\pa} formula.

\paragraph{Synchronize the elimination of exponential occurrences of the same variable in different atomic formulas}

Although Lemma~\ref{lem:exp-ineq} is stated for a single atomic formula, the elimination of the exponential occurrences of the same variable in different atomic formulas can actually be synchronized. That is,  let $\alpha^\tau_{1}, \alpha^\tau_{2}, B^\tau, \delta^\tau$ be the constants as stated in the aforementioned under-approximation of an inequality $\tau$, define $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ as the minimum of $\alpha^\tau_1$, the maximum of $\alpha^\tau_2$, the maximum of $B^\tau$, and the maximum of $\delta^\tau$ respectively with $\tau$ ranging over the inequalities of $\varphi$. Then we can use the same constants $\alpha^{\min}_1, \alpha^{\max}_2, B^{\max}, \delta^{\max}$ for different inequalities when eliminating the exponential occurrences of the same variable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
Note that when Lemma~\ref{lem:exp-ineq} is applied to an atomic formula $\tau$ to eliminate exponential occurrences of $\ivarx$, it computes 4 parameters, namely $(\alpha^\tau_1,\alpha^\tau_2,B^\tau,\delta^\tau)$, and obtains an equivalent {\pa} formula according to these parameters. Since the parameters are used as boundary conditions in enumeration, we can choose the strictest boundary conditions and apply to all atomic formulas.

@@

For instance, assume coefficients of $10^\ivarx$ in atomic formulas of $\varphi$ are all non-negative. Let $\varphi^+$ denote the formula obtained by replace atomic formulas where coefficients of $10^\ivarx$ are positive by $\ltrue$. The resulted formula $\varphi'$ is
\[
\small
\begin{array}{l}
\bigvee \limits_{p=0}^{B^{\max}-1} ( \ivarx_{\sigma(i-1)} = p \wedge \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)}) [p/\ivarx_{\sigma(i-1)}])   \\
\bigvee \left(
    \begin{array}{l}
        \ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  - 1  \wedge \\
        \ivarx_{\sigma(i-1)} \ge \ivarx_{\sigma(i-2)} +\delta^{\max} \wedge \varphi^+  
    \end{array}
    \right)\\
%
\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \le \alpha^{\min}_1  -1 \ \wedge \\
 \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right)\\
\bigvee \bigvee \limits_{p=0}^{\delta^{\max}-1} 
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} \ge \alpha^{\max}_2 + 2 \ \wedge \\
 \ivarx_{\sigma(i-1)} = \ivarx_{\sigma(i-2)} +p\ \wedge\\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[\ivarx_{\sigma(i-2)} + p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right)\\
\bigvee \bigvee \limits_{p = \alpha^{\min}_1}^{\alpha^{\max}_2 + 1}
\left(
\begin{array}{l}
\ivarx_{\sigma(i-1)} \ge B^{\max} \wedge \ivarx_{\sigma(i-1)} = p\ \wedge \\
 \varphi(\ivarx_{\sigma(i)}, \ldots, \ivarx_{\sigma(1)})[p /\ivarx_{\sigma(i-1)}] 
\end{array}
\right),
\end{array}
\]
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The {\pa} formula resulting from the elimination of exponential occurrences is essentially a big disjunction of the formulas of small size. If we store this big disjunction naively, then the formula size quickly blows up and exhausts the memory. Instead, we choose to do a depth-first search (DFS) and consider the disjuncts, which are of small sizes, one by one, and solve the satisfiability problem for these disjuncts. If during the search, a satisfiable disjunct is found, then the search terminates and ``SAT'' is reported.

\paragraph{Preprocess with small upper bound}
